\subsection*{1/24}

Meeting minutes ended up just being a to-do list. (Make a wiki, make a
git repository, everyone read Sam's notes, I’ll create pltphb@gmail
and start a calendar, we need a fifth group member).  

\subsection*{1/30}

Orders of business:
\begin{enumerate}
\item Meetings are Sundays, 8:30, Sam's suite. 
\item Everybody should read Sam's modeling papers for next week and
  send her whiny emails if they don't make sense.
\item Everybody should also read the git tutorial that Will is going
  to send/post on the wiki.
\end{enumerate}

Also:
\begin{itemize}
\item Will, do you want to just make everyone wiki accounts now so we
  don't have to deal with it later?  
\item Rob: Eva had a good point about how I don't really need the
  calendars anymore (or at least for a while) now that we have a set
  meeting time, so you don't have to do whatever you were going to try
  to make Google Calendar behave properly.
\end{itemize}

\subsection*{2/6}

\begin{enumerate}
\item Sam's materials and explanation. We spent the most time on:
  \begin{itemize}
  \item concept: time-stepping
  \item time-stepping as a concept
  \item methods for solving diff eqs
    \begin{itemize}
    \item Euler, implicit, explicit, semi-implicit
    \item stability/error-checking
    \end{itemize}
  \item forces we're using
  \end{itemize}

\item Our language is for physicists who can't program.
\item Mission: Language Spec now underway. Meet with Aho in a week to
  get his input

\item Are we doing collisions? Conclusion: probably
\item Post-Sam: Will's rendering rant
\end{enumerate}


\subsubsection*{Project Milestones}
\begin{enumerate}
\item Finish language spec in a week
  \begin{itemize}
  \item show it to Aho
  \item whitepaper 1.5 wks after the meeting
  \item Will's putting his ideas up on a page, then we'll use the wiki
    talk feature to add and modify things
  \end{itemize}

\item Rendering:
  \begin{itemize}
  \item OpenGL for Java
  \item display single triangles
  \item mesh framework for triangles
  \end{itemize}

\item Semi-implicit Euler and spring forces (followed by more forces)
\item Implicit time stepping
\item discrete geometry
\item collisions (in parallel with discrete geometry)
\end{enumerate}

\subsection*{2/20}

Tonight we write the whitepaper!

\subsection*{2/27}

\subsubsection*{Things We Are Going To Code}
(note: this outline does not imply order)
\begin{enumerate}
\item Rendering \textbf{(Rob)}
  \begin{itemize}
  \item file output
  \item video (onscreen)
  \item displaying triangles
  \item updating
  \item aesthetics \textbf{(Sam)}
  \end{itemize}

\item Time-stepping functions \textbf{(Colleen)}
  \begin{itemize}
  \item semi-implicit
  \item implicit
  \end{itemize}

\item Forces \textbf{(Sam)}

\item Discretizing geometry \textbf{(Will and Rob)}
\item Collisions \textbf{(Eva)}
\item Translator \textbf{(Will)}
  \begin{itemize}
  \item define API for target source code (see Will's outline in lang
    specs for ideas) 
  \item yaaay regexp
  \end{itemize}

\item UNIT TEST EVERYWHERE \textbf{(Eva)}
\end{enumerate}

\subsubsection*{Gameplan}
Rendering and the translator are independent of other things.

\subsection*{3/21}

\subsubsection*{What to do on the manual}
\begin{description}
\item[Missing] sections on builder syntax camera placement, and
  discrete geometry. The rest is filling in TODOs and 2.6.
\item[Dividing things up] write about the section you're
  implementing. Will has to do lots of things because he understands
  them! Everything else is first-come-first-serve.
\item[The plan] tasks:
  \begin{itemize}
  \item Will finishes tutorial outline by Friday night
  \item Everyone picks their sections by Saturday at midnight, or I
    assign them. Shoot for even distribution of work, obviously.
  \item CHECK YOUR EMAILS PLEASE
  \item Hard deadline for submitting content to Will is Tues, 3/29, at 10:30am.
  \item No meeting this Sunday, since it doesn't seem like we need it.
  \end{itemize}
\end{description}

\subsection*{4/3}

\subsubsection*{Progress reports}
\begin{description}
\item[Will] Translator for most of the Feynstein-specific syntax is
  implemented. Still working on the Java parts (i.e. telling
  translator when it should ignore things that are Java). ~500 lines
  of C, will probably be close to double that.

\item[Rob] Having issues with JOGL implementation--code not might be
  correct. (Code: rendering triangles to files, video out.)
\end{description}

\subsubsection*{How to get things done better}
\begin{itemize}
\item Deadlines:
  \begin{itemize}
  \item Content deadline: 96 hours before due date
  \item Proofing deadline: 72 hours before due date
  \item Changes-from-proofing deadline: 48 hours before
  \item Reproofing deadline: 24 hours before
  \end{itemize}
\item MY BRILLIANT PLAN:
  \begin{itemize}
  \item If your material is late, a note is made on a new wiki page
    (will be created soon) and another person takes the material. If
    you take over someone else's late material, a note is made on the
    wiki page about how awesome you are.
  \item Meetings may involve reassigning work
  \item For each note, we suggest Aho lower the person's personal
    project grade a letter.
  \end{itemize}
\end{itemize}

\subsubsection*{Deadlines for EVERYTHING}
\begin{enumerate}
\item Rendering (Rob): done by Thurs, April 7 (night)
  \begin{itemize}
  \item file output
  \item video (onscreen)
  \item displaying triangles
  \item updating
  \item aesthetics (Sam)
  \end{itemize}

\item Translator (Will): done by Sunday, April 10

\item Obj importer (Sam): done by Sunday, April 10
\item Gravity (Sam): done by Thursday, April 14
\item Shape primitives (Will and Rob): Sunday, April 17
\item Time-stepping functions (Colleen plus Sam and Eva): done by Sunday, April 17
semi-implicit (first)
\item Forces (Sam): done by Sunday, April 24th
\item Time-stepping II (Colleen and Sam): done by Sunday, May 1
\item Keyboard and mouse navigation (Rob): done by Sunday, May 1
\item Collisions (Eva): done by Sunday, May 1
\item UNIT TEST EVERYWHERE (Eva): Testing must be done by Sunday, May 8
\item Textures, shading, etc (Systems people): as needed
\item Project report: Friday, April 29
\end{enumerate}

\subsection*{4/10}

\subsubsection*{How to git}
One person working on things at a time, otherwise problems.  This
means check with everyone before you commit!

\subsubsection*{Other things}
\begin{itemize}
\item Automated compilation
  \begin{itemize}
  \item Command: run
  \item shell script in top-level directory
  \item can only compile things inside git directory
  \end{itemize}
\item temp file: put temp stuff there (e.g. tests)
\item Before you commit code, run make clean in the top level
  directory (to get rid of annoying .class files)
\end{itemize}

\subsubsection*{Stuff due soon}
\begin{itemize}
\item Systems people: object stuff
\item time stepping (Sam and Colleen meeting Friday)
\item Will: by end-of-night Sunday, implement main method using
  renderer and FileShape class.
\end{itemize}

\subsection*{4/18}

\subsubsection*{Will's updates}
\begin{itemize}
\item API changes: eg forces none;
\item shapes addable; shapes stored in right place; close to storing meshes
\item still need to test
\end{itemize}

\subsubsection*{Sam's updates}
\begin{itemize}
\item Renderer (w/timestepping) is running in a main method in
  java--not Feynstein yet. Working on that now.
\item Set up to "do physics instead of hard coding"
\item going to work on cleaning up API
\item Note to everyone: EVERYTHING EXTENDS SCENE. (Scene is abstract.)
\item some things are already defined in abstract class, e.g. stepping
  through forces physics, etc
\end{itemize}

\subsubsection*{Colleen's updates}
\begin{itemize}
\item Please be better about responding to my emails.
\item Demo date: May 11th, 2:30pm
\item Meeting with Aho: try for this Thursday, 12-2?
\end{itemize}

\subsubsection*{Revising of deadlines, addition of missing stuff}
(see master deadlines page)

\subsection*{4/24}

\subsubsection*{Things we did}
\begin{itemize}
\item Will: meshes!
\item Sam: gravity! Spring force!
\end{itemize}

\subsubsection*{Deadline musical chairs again}

\subsubsection*{Other stuff}
our final test: www.cs.ubc.ca/~rbridson/movies/spinningcloth.mov
(this depends on Rob learning about textures)

\subsection*{5/3}

\subsubsection*{There are problems.}
\begin{itemize}
\item Setting a deadline about 24 hours after the end of our final was a Bad Idea. Leniency is granted for missed deadlines in light of that fact.
\item Nobody is finished with anything
\item Things people have done:
  \begin{itemize}
  \item Will: a bunch of collisions stuff
  \item Sam: all the forces, stressing over implicit time stepping
  \item Eva: spring force tests
  \item Rob: mapping java to feynstein code (for error handling)
  \item Colleen: stressing over implicit time stepping
  \end{itemize}
\item Eva and Rob will receive phone calls, not emails, about
  scheduling from now on
\end{itemize}

\subsubsection*{When things can be done}
\begin{itemize}
\item Rob: error handling due by 6pm tomorrow
\item Eva: translator tests due by 3pm, force tests by 6pm
\item Colleen: collisions: proximity tests hopefully done tonight,
  CTCD due...tomorrow?
\item Sam: don't forget about file output and mouse navigation (not
  that we'll let you), but focus on whatever is most important at the
  moment
\end{itemize}