\input{/home/haldean/Documents/latex_template.tex}
\title{Team Seven Project Report: Feynstein}
\subtitle{A Language for Physical Simulation}

\author{Samantha Ainsley \and Eva Asplund \and William Brown \and
  Colleen McKenzie \and Robert Post}

\newtoggle{tutorial}
\newtoggle{manual}
\newtoggle{listing}

\toggletrue{tutorial}
\toggletrue{manual}
\toggletrue{listing}

\begin{document}
\maketitle

\begin{abstract}
  Feynstein is a language designed for making physical simulation
  accessible to those who want to experiment with physics but who do not
  have significant experience in computer science. Feynstein will make
  the process of going from an experiment’s conception to its simulation
  simple and fast, and will provide the end user with accurate video
  renderings, either in a file or on screen.
\end{abstract}

\tableofcontents

\section{Introduction}

Historically, if users wanted to have the power of a full physical
simulation engine, they had to either purchase and learn very complex
and expensive 3D simulation software, or write it themselves (usually
in C++). Because these options are costly and time-consuming, often
students of physics will learn simulation using a software package
like Matlab or Mathematica. These tools have the benefit of
abstracting much of the complexity of the task, but with the loss of
complexity comes a loss of control. Feynstein attempts to be a
midpoint between these two extremes; it is a language that allows for
the full expressivity of a complex simulation package, but with the
ease-of-use that comes with a tool like Matlab.

Feynstein's native simulator and its simple and convenient syntax make
it an ideal tool for researchers with limited programming experience
for configuration simulations to test experimental
hypothesis. Moreover, the language offers an excellent framework for
teaching the fundamentals of physics-based computer simulation to
students with limited understanding of mechanics or programming.

Feynstein is fully object oriented, allowing for easy
extensibility. Users can customize integration and rendering methods,
giving them full control of how their animation is calculated and
displayed. Finally, users can use the built-in forces and 3D solids
and define their own. Feynstein is appropriate for all levels of user
expertise, whether they want to simulate a sphere on a spring or a
complex compression force on a custom shape.

\iftoggle{tutorial}{
\section{Language Tutorial}
\input{tutorial.tex}
}{}

\iftoggle{manual}{
\section{Language Reference Manual}
\input{manual.tex}
}{}

\section{Project Plan}
\emph{Colleen McKenzie, Project Manager}

\input{colleen.tex}

\section{Language Evolution}
\emph{Samantha Ainsley, Language Guru} \\

\input{sam.tex}

\section{Language Architecture}
\emph{William Brown, System Architect} \\

\input{will.tex}

\section{Development Tools}
\emph{William Brown, System Architect}

\input{rob.tex}

\section{Testing Plan}
\emph{Eva Asplund, System Tester} \\

\input{eva.tex}

\section{Conclusion}
\input{lessons.tex}

\startappendix

\section{Project Contributions}
\label{sec:whowhat}
\input{whowhat.appendix.tex}

\section{Project Log}
\label{sec:projectlog}
\input{log.appendix.tex}

\section{Source Code Listing}
\label{sec:listing}
\iftoggle{listing}{
  \input{feynstein.listing.tex}
  \input{tests.listing.tex}
  \input{translator.listing.tex}}{}

\end{document}